Kronos \& Krios: Evidence for accretion of massive, rocky planetary system in a sun-like pair

### 2017-06-06 paper/ms.tex
Work on comments on data section.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{spec2.pdf}
  \caption{Same as \figname~\ref{fig:spec1}
    but for smaller portions of spectra at longer wavelengths that are
    not dominated by \elem{Fe}.
    We mark elements that give rise to strong lines.
    The feature in the third column of the second row is a strong \elem{Li}
    line that was not modeled by \citealt{2016ApJS..225...32B}. This line was
    studied in a separate work (\citealt{jmlithium}).
    Note that the lines of \elem{Na} and \elem{O} which are under-enhanced
    in \bizarreone\ relative to \elem{Fe} or other refractory elements
    show weaker residuals.
  }
  \label{fig:spec2}
\end{figure}

\sunanalog\ and \bizarreone\ were identified as a
candidate comoving star pair in our recent search for comoving stars using the
proper motions and parallaxes from the \tgas catalog, a component of \gaia\ \dr\
(the astrometric measurements are listed in \tablename~\ref{tab:kk}).
We refer the readers to this previous work (\citealt{2016arXiv161202440O}) for a
full explanation of the methodology behind this search and only briefly describe
the method here.
For a given pair, we compute a marginalized likelihood ratio between the
hypotheses (1) that a given pair of stars share the same 3D velocity vector, and
(2) that they have independent 3D velocity vectors, both given only observations
of two components of the velocities (proper motions).
We then select a sample of high-confidence comoving star pairs by making a
conservative cut on this likelihood ratio.
In the resulting catalog of comoving pairs (\citealt{2016arXiv161202440O}),
the pair presented in this paper was assigned a group id of 1199,
and the marginalized likelihood ratio (Bayes factor)
between the two hypotheses is $\ln{\mathcal{L}_1/\mathcal{L}_2} = 8.52$,
well above the adopted cut value of 6.
The pair has also been previously recognized as a visual double star system
in Washington Double Star catalog (\citealt{2001AJ....122.3466M}).
We have checked that we do not find any possible additional comoving companions
by lowering the likelihood ratio cut for the stars around this pair.

In a separate effort to study detailed chemical abundances of potential
planet-hosting stars, \citet{2016ApJS..225...32B} analyzed high resolution
spectra of both stars obtained using HIRES spectrograph on Keck~I telescope.
The spectral resolution is $R\approx 70000$ and the wavelength coverage is
$5164-7799$~\AA.
Typical signal-to-noise ratio in continuum is $>200$~per pixel.
The resulting measurements include abundances measured for 15 chemical species
(C, N, O, Na, Mg, Al, Si, Ca, Ti, V, Cr, Mn, Fe, Ni, Y) as well as stellar parameters
and high precision radial velocities.
For the details of the spectral analysis, we refer the readers to
\citealt{2016ApJS..225...32B}.
The \elem{Li} doublet at $6707.6$~\AA\ for this sample was
investigated in a separate work (\citealt{jmlithium}).
We list all spectroscopic measurements including
the absolute abundance\footnote{
  Absolute abundance of element \elem{X} is $A(\elem{X}) = \log_{10} (n_\elem{X}/n_\elem{H})$
  where $n_\elem{X}$ is the number density of element \elem{X}.
} of \elem{Li} for the two stars in Table~\ref{tab:kk}.

Although selected based only on their proper motions, the two stars
have identical radial velocities within uncertainties (Table~\ref{tab:kk}),
confirming that they are comoving in all directions.
The projected separation between the pair is 1.9\arcmin\ ($\approx 0.01$~pc),
and the 3D separation is $\approx 0.6$~pc.
Given their proximity in phase space, it is generally accepted to assume that
the two stars were born together perhaps to form a wide binary system from the
same birth cloud.
Thus, we expect the two stars to have identical metallicities and abundance patterns.
Surprisingly, however, one of the stars, \bizarreone\ is significantly more metal
rich than the other by 0.2~dex ($\approx 60\%$; \figname~\ref{fig:abundances}).
What is more puzzling is that not all elements are equally enhanced:
the abundances of \bizarreone\ shows selective depletion in
\elem{C}, \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}
relative to \elem{Fe}.

The validity of the measured abundance differences is further demonstrated
in \figname~\ref{fig:spec1} and \ref{fig:spec2} where we show
segments of the spectra and model of the two stars.
As expected from metallicity difference ($\Delta\feh \approx 0.2$),
the ratio of data and model between the two stars show significant
residuals of almost all metal features, largely dominated by \elem{Fe}.
However, for lines of elements that are not as enhanced in \bizarreone\,
the residuals are much smaller in amplitude.
%Li?

We stress that none of the other four twin-like ($\Delta T_\mathrm{eff}
\lesssim 100$~K) wide binary pairs examined in \citealt{2016ApJS..225...32B}
show discrepancies in abundances between the stars at this level.
As shown in \figname~\ref{fig:deltaXH},
the differences in other pairs for all elements except \elem{N} and \elem{O},
which are also the most uncertain (\tablename~\ref{tab:kk})
are less than $0.05$~dex, makiing \bizarreone-\sunanalog\ pair an outlier.
Moreoever, the abundance differences cannot be due to different chemical enrichment
history in the disk even if the two stars were of different origin.
\figname~\ref{fig:deltaXH} additioanlly shows the distribution of $\Delta[\elem{X}/\elem{H}]$
for random pairings of two stars with similar \feh\ as \bizarreone\ and \sunanalog\
as violins for each element.
We see that when a star is enhanced in $\elem{Fe}$ by $0.2$~dex,
all other elements are enhanced at a similar level with some variations.
Specifically, for a typical star with $\elem{Fe} \approx 0.2$~dex,
we generally expect $[\elem{Na}/\elem{Fe}] > 0$ and $[\elem{Mn}/\elem{Fe}] > -0.1$
(\citealt{Battistini:2015aa,Bensby:2003aa})
making the low $\elem{Na}/\elem{Fe}$ and $\elem{Mn}/\elem{Fe}$ seen in \bizarreone\
very unlikely to arise from variations in Galactic chemical evolution.

@@

The fact that the two stars share a very similar space motion
at a separation of $\approx 0.6$~pc strongly argues that the pair is coeval.
We consider other age indicators apart from their phase space coordinates
to assess the ages and coevality of the two stars.
First, given the precise measurements of $\log(g)$, $T_\mathrm{eff}$,
comparing these values to theoretical isochrones may also be indicative of the two stars' ages.
We perform isochrone fitting using the Yale-Yonsei model isochrones (\citealt{2013ApJ...776...87S}).
The input data are $\log(g)$, $T_\mathrm{eff}$, \feh, parallaxes, $B$-band magnitudes, and their errors.
The isochrone ages of \sunanalog\ and \bizarreone\ are
$4.00_{-1.56}^{+1.51}$~Gyr and $4.28_{-1.03}^{+1.11}$~Gyr, respectively,
consistent with them being coeval.

Surface lithium abundance in a sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning.
Thus, surface lithium abundance is an indicator of stellar ages.
The $\elem{Li}$ abundance of $2.25$ for \sunanalog\ implies an age of $\lesssim
1$~Gyr according to the theoretical models tuned to explain the solar lithium
abundance and rotational profile (\citealt{2005Sci...309.2189C}).
The lithium abundance of \bizarreone\ is still higher by $0.5$~dex showing the
largest difference among all measured elements.
This translates to $\sim 500$~Myr difference in age.
Given the overall higher metal abundances and the peculiar abundance patterns
in \bizarreone, it is unclear, however, whether this higher $\elem{Li}$
abundance means a younger age or something else.
For example, the presence of $\elem{Li}$-rich red giant stars has been
attributed to the engulfment of substellar companions such as gas giant planets
or brown dwarfs which may replenish $\elem{Li}$ (\citealt{Casey:2016aa}).

In fact, the surface lithium abundance is the only indication that points to younger ages.
If the two stars are indeed only several hundred Myrs old at most,
they are expected to be part of a larger comoving group of stars.
However, as we mention above, there is no evidence in our search of comoving pairs
using \tgas\ that the two stars belong to a larger group of young stars.
Very young stars are often show signs of activity such as
X-ray emission from magnetic activity, emission lines, or infrared excess due to
circumstellar disks.
We have looked for these signs for the pair to find none, and their spectral energy
distributions are consistent with blackbody.
The low $v\sin(i)$ values also argue against very young ages inferred from the
surface lithium abundances.
% TODO: reference for vertical action as age?
Finally, the pair's fiducial orbit in a axisymmetric Milky Way-like potential
has a vertical action larger than the Sun, favoring an older age.
We concluded that the two stars are most likely coeval, $\sim 4$~Gyr old main
sequence stars, and that their unusually high \elem{Li} abundance requires an
alternative explanation.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{TcRank_deltaXH_concise.pdf}
  \caption{Abundance difference of \bizarreone-\sunanalog\ pair
    ranked by the condensation temperature of elements for solar composition gas
    from \citealt{2003ApJ...591.1220L}.
    The condensation temperature may be read from the gray line and right y-axis.
    We show three wide binary systems selected from the literature:
    HD~20782/1 (\citealt{Mack:2014aa}, $\feh\approx0$),
    XO-2N/S (\citealt{Biazzo:2015aa}, $\feh\approx0.35$),
    and WASP-94AB (\citealt{Teske:2016aa}, $\feh\approx0.3$).
    Locations of elements with at least one measurement from any study
    are indicated by a vertical line and its symbol.
    Note that often multiple values are reported for one element corresponding
    to different ionization states in equivalent width analyses.
    No other pair studied so far were shown to have such large difference
    in metallicity or sharp contrast between (moderately) volatile and
    refractory elements as \bizarreone-\sunanalog.
  }
  \label{fig:relabun_tcrank}
\end{figure}

>>>>

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{spec2.pdf}
  \caption{Same as \figname~\ref{fig:spec1}
    but for smaller portions of spectra at longer wavelengths that are
    not dominated by \elem{Fe}.
    We mark elements that give rise to strong absorption lines.
    The feature in the third column of the second row is a strong \elem{Li}
    line that was not modeled by \citealt{2016ApJS..225...32B}; this line was
    studied in a separate work (\citealt{jmlithium}).
    Note that the lines of \elem{Na} and \elem{O}, which are under-enhanced
    in \bizarreone\ relative to \elem{Fe} or other refractory elements,
    show weaker residuals.
  }
  \label{fig:spec2}
\end{figure}

\sunanalog\ and \bizarreone\ were identified as a
candidate comoving star pair in our recent search for comoving stars using the
proper motions and parallaxes from the \tgas\ catalog, a component of \gaia\ \dr\
(the astrometric measurements are listed in \tablename~\ref{tab:kk}).
We refer the readers to this previous work (\citealt{2016arXiv161202440O}) for a
full explanation of the methodology behind this search and only briefly describe
the method here.
For a given pair, we compute a marginalized likelihood ratio between the
hypotheses (1) that a given pair of stars share the same 3D velocity vector, and
(2) that they have independent 3D velocity vectors, both given only observations
of two components of the velocities (parallaxes and proper motions).
We then select a sample of high-confidence comoving star pairs by making a
conservative cut on this likelihood ratio.
In the resulting catalog of comoving pairs (\citealt{2016arXiv161202440O}),
the pair presented in this paper was assigned a group id of 1199,
and the marginalized likelihood ratio (Bayes factor)
between the two hypotheses is $\ln{\mathcal{L}_1/\mathcal{L}_2} = 8.52$,
well above the adopted cut value of 6.
The pair has also been previously recognized as a visual double star system
in Washington Double Star catalog (\citealt{2001AJ....122.3466M}).
We have checked that we do not find any possible additional comoving companions
by lowering the likelihood ratio cut for the stars around this pair.

In a separate effort to study detailed chemical abundances of potential
planet-hosting stars, \citet{2016ApJS..225...32B} obtained and analyzed high resolution
spectra of both stars using the HIRES spectrograph on the Keck~I telescope.
The spectral resolution is $R\approx 70000$ and the wavelength coverage is
$5164--7799$~\AA.
A typical signal-to-noise ratio in the spectral continuum is $>200$~per pixel.
The resulting measurements include elemental abundances for 15 chemical species
(C, N, O, Na, Mg, Al, Si, Ca, Ti, V, Cr, Mn, Fe, Ni, Y) as well as stellar parameters
and high precision radial velocities.
For the details of the spectral analysis, we refer the readers to
\citealt{2016ApJS..225...32B}.
The \elem{Li} doublet at $6707.6$~\AA\ --- clearly visible in both spectra (see
\figname~\ref{fig:spec2}) --- for this sample was investigated in a separate
work (\citealt{jmlithium}).
We list all spectroscopic measurements including
the absolute abundance\footnote{
  Absolute abundance of element \elem{X} is
  $A(\elem{X}) = \log_{10} (n_\elem{X}/n_\elem{H})$
  where $n_\elem{X}$ is the number density of element \elem{X}.
}
of \elem{Li} for the two stars in Table~\ref{tab:kk}.

Although selected based only on their astrometry, the two stars
have identical radial velocities within uncertainties (Table~\ref{tab:kk}),
confirming that they are truly comoving.
The projected separation between the pair is 1.9\arcmin\ ($\approx 0.01$~pc),
and the 3D separation is $\approx 0.6$~pc.
% note on uncertainties
Given their proximity in phase space, it is generally accepted to assume that
the two stars were born together and, perhaps, are a bound wide binary system.
Thus, we expect the two stars to have identical metallicities and abundance patterns.
Surprisingly, one of the stars, \bizarreone\ is significantly more metal
rich than the other (by 0.2~dex $\approx 60\%$; \figname~\ref{fig:abundances}).
Even more puzzling is that not all elements are equally enhanced:
the abundances of \bizarreone\ show selective depletion in
\elem{C}, \elem{N}, \elem{O}, \elem{Na}, and \elem{Mn}
relative to \elem{Fe}.

The validity of the measured abundance differences is further demonstrated
in \figname~\ref{fig:spec1} and \ref{fig:spec2} where we show
segments of the spectra and models of the two stars
used to measure their abundances.
As expected from their reported metallicity difference ($\Delta\feh \approx 0.2$),
the ratio of data and model between the two stars show significant
residuals of almost all metal line features, largely dominated by \elem{Fe}.
However, for lines of elements that are not as enhanced in \bizarreone\,
the residuals are much smaller in amplitude.
%Li?

We stress that none of the other four twin-like ($\Delta T_\mathrm{eff}
\lesssim 100$~K) wide binary pairs examined in \citealt{2016ApJS..225...32B}
show discrepancies in abundances between the stars at this level.
As shown in \figname~\ref{fig:deltaXH},
the differences in other pairs for all elements except \elem{N} and \elem{O},
which are also the most uncertain (\tablename~\ref{tab:kk}),
are less than $0.05$~dex, making \bizarreone-\sunanalog\ pair a significant outlier.

The abundance differences cannot be due to different chemical enrichment
histories in the disk even if the two stars were of different origin.
\figname~\ref{fig:deltaXH} additionally shows the distribution of abundance
differences, $\Delta[\elem{X}/\elem{H}]$, between random pairings of two stars
with similar \feh\ as \bizarreone\ and \sunanalog\.
We see that when a star is enhanced in $\elem{Fe}$ by $0.2$~dex,
all other elements are typically enhanced at a similar level, with some variations.
Specifically, for a typical star with $\feh \approx 0.2$~dex,
we generally expect $[\elem{Na}/\elem{Fe}] > 0$ and $[\elem{Mn}/\elem{Fe}] > -0.1$
(\citealt{Battistini:2015aa,Bensby:2003aa})
making the low [\elem{Na}/\elem{Fe}] and [\elem{Mn}/\elem{Fe}] seen in \bizarreone\
very unlikely to arise from variations in Galactic chemical evolution.

@@

The fact that the two stars share a nearly identical space motion
at a separation of $< 1$~pc strongly argues that the pair is coeval.
We therefore  consider other age indicators apart from their phase space coordinates
to assess the ages and coevality of the two stars.
First, given the precise measurements of $\log(g)$ and $T_\mathrm{eff}$,
we can constrain the ages of the two stars by
comparing these values to theoretical isochrones.
We perform isochrone fitting to these values using the Yale-Yonsei model
isochrones (\citealt{2013ApJ...776...87S}).
The input data are $\log(g)$, $T_\mathrm{eff}$, \feh,
parallaxes, $B$-band magnitudes, and their errors.
The best-fit isochrone ages of \sunanalog\ and \bizarreone\ are
$4.00_{-1.56}^{+1.51}$~Gyr and $4.28_{-1.03}^{+1.11}$~Gyr, respectively,
consistent with them being coeval.

The Surface lithium abundance in a sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning.
Thus, surface lithium abundance is an indicator of stellar ages.
The absolute $\elem{Li}$ abundance of $2.25$~dex for \sunanalog\ implies an age
of $\lesssim 1$~Gyr according to the theoretical models tuned to explain the
solar lithium abundance and rotational profile (\citealt{2005Sci...309.2189C}).
The lithium abundance of \bizarreone\, $2.74$~dex, shows the largest difference
among all measured elements.
This translates to a $\sim 500$~Myr difference in age.
Given the overall higher metal abundances and the peculiar abundance patterns
in \bizarreone, it is unclear, however, whether this higher $\elem{Li}$
abundance means a younger age or something else.
For example, the presence of $\elem{Li}$-rich red giant stars has been
attributed to the engulfment of substellar companions such as gas giant planets
or brown dwarfs which may replenish $\elem{Li}$ (\citealt{Casey:2016aa}).

In fact, the surface lithium abundance is the only indication that points to younger ages.
If the two stars are indeed only several hundred Myrs old at most,
they are expected to be part of a larger comoving group of stars.
However, as we mention above, there is no evidence in our search of comoving pairs
using \tgas\ that the two stars belong to a larger group of young stars.
Very young stars are often show signs of activity such as
X-ray emission from magnetic activity, emission lines, or infrared excess due to
circumstellar disks.
We have looked for these signs for the pair to find none, and their spectral energy
distributions are consistent with blackbody.
The low $v\sin(i)$ values (\tablename~\ref{tab:kk}) also argue against very
young ages inferred from the surface lithium abundances.
% TODO: reference for vertical action as age?
Finally, the pair's fiducial orbit in a axisymmetric Milky Way-like potential
has a vertical action larger than the Sun, favoring an older age.
We therefore conclude that the two stars are most likely coeval, $\sim 4$~Gyr old main
sequence stars, and that their unusually high \elem{Li} abundance requires an
alternative explanation.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{TcRank_deltaXH_concise.pdf}
  \caption{Abundance differences of the \bizarreone-\sunanalog\ pair
    ranked by the condensation temperature of elements for solar composition gas
    from \citealt{2003ApJ...591.1220L}.
    The condensation temperature may be read from the gray line and right y-axis.
    We show three wide binary systems selected from the literature:
    HD~20782/1 (\citealt{Mack:2014aa}, $\feh\approx0$),
    XO-2N/S (\citealt{Biazzo:2015aa}, $\feh\approx0.35$),
    and WASP-94AB (\citealt{Teske:2016aa}, $\feh\approx0.3$).
    Locations of elements with at least one measurement from any study
    are indicated by a vertical line and its symbol.
    Note that often multiple values are reported for one element corresponding
    to different ionization states in equivalent width analyses.
    No other pair studied so far were shown to have such large difference
    in metallicity or sharp contrast between (moderately) volatile and
    refractory elements as \bizarreone-\sunanalog.
  }
  \label{fig:relabun_tcrank}
\end{figure}
